sdmTMB: an R package for fast, flexible, and user-friendly generalized linear mixed effects models with spatial and spatiotemporal random fields

### 2023-03-11 doc/paper-jss/inla-comparison.Rmd
more \pkg #13

# Comparison between INLA, inlabru, and sdmTMB {#app:inla}

@@

In INLA, a PC-prior can be used for the temporal autoregressive parameter, i.e., a prior that defines the probability the autocorrelation parameter is greater than a given value (e.g., `P(cor > 0) = 0.9`), but for now, sdmTMB only takes a Gaussian prior (e.g., `priors = sdmTMBpriors(ar1_rho = normal(0, 1))`).

@@

## INLA version

@@

## inlabru version

@@

## sdmTMB version

@@

```{r maps, out.width="5.5in", fig.cap="Predicted occurence probabilities in time and space from the sdmTMB model.", fig.align='center'}
inverse_link <- m$family$linkinv
ggplot(pred) +
  geom_raster(aes(X, Y, fill = inverse_link(est))) +
  facet_wrap(~year) +
  scale_fill_viridis_c() +
  coord_fixed()
```

@@

```{r inlabru-pred, out.width="5.5in", fig.cap="Predicted occurence probabilities in time and space from the inlabru model.", fig.align='center'}
nd_bru <- nd
nd_bru <- dplyr::left_join(nd_bru, yr_lu, by = "year")
sp::coordinates(nd_bru) <- c("X", "Y")
nd_bru <- as(nd_bru, "SpatialPixelsDataFrame")
pred_bru <- predict(m_bru, data = nd_bru, formula = ~ plogis(f + fac))

>>>>

# Comparison between \pkg{INLA}, \pkg{inlabru}, and \pkg{sdmTMB} {#app:inla}

@@

In INLA, a PC-prior can be used for the temporal autoregressive parameter, i.e., a prior that defines the probability the autocorrelation parameter is greater than a given value (e.g., `P(cor > 0) = 0.9`), but for now, \pkg{sdmTMB} only takes a Gaussian prior (e.g., `priors = sdmTMBpriors(ar1_rho = normal(0, 1))`).

@@

## \pkg{INLA} version

@@

## \pkg{inlabru} version

@@

## \pkg{sdmTMB} version

@@

```{r maps, out.width="5.5in", fig.cap="Predicted occurence probabilities in time and space from the \pkg{sdmTMB} model.", fig.align='center'}
inverse_link <- m$family$linkinv
ggplot(pred) +
  geom_raster(aes(X, Y, fill = inverse_link(est))) +
  facet_wrap(~year) +
  scale_fill_viridis_c() +
  coord_fixed()
```

@@

```{r inlabru-pred, out.width="5.5in", fig.cap="Predicted occurence probabilities in time and space from the \pkg{inlabru} model.", fig.align='center'}
nd_bru <- nd
nd_bru <- dplyr::left_join(nd_bru, yr_lu, by = "year")
sp::coordinates(nd_bru) <- c("X", "Y")
nd_bru <- as(nd_bru, "SpatialPixelsDataFrame")
pred_bru <- predict(m_bru, data = nd_bru, formula = ~ plogis(f + fac))
